---
title: "Reproducible Document"
author: "Oliver Bonde van Zwol"
format: html
---

```{r setup}
#| message: false
#| warning: false

library("tidyverse")

post_meal_data <- read.csv2(here::here("data/post-meal-insulin.csv"))


small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi) # make a new df with only the selected columns


tidier_post_meal_data <- small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) %>%
  filter(glykemi == 0) %>%
  select(-glykemi)
```

## About me

-   Oliver
-   PhD

Just started on a **PhD** at *FGM*

## Simple code

```{r}
3 * 3
```




## Writing R code

2+2

```{r}
2 + 2
```

## Testing for Git

It is really cold outside


## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continous variable

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

See @fig-bmi-histo


## Exercise: discrete plots

```{r}
#| fig-cap: "Groups"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```


## Side by side plot
```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```


## Selecting  data

```{r}
select(post_meal_data, Age) # Selects only the Age column

select(post_meal_data, Age, BMI, Weight) # select only Age, BMI, and Weight column

select(post_meal_data, -Group) # removes the Group columns

select(post_meal_data, starts_with("PG")) # selects only columns starting with PG

select(post_meal_data, ends_with("DXA")) # selects all columns ending with DXA

select(post_meal_data, contains("0")) # selects all columns containing 0
```


## Renaming columns

```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```



## Filtering data


```{r}
small_post_meal_data %>%
  filter(glykemi == 1)

small_post_meal_data %>%
  filter(BMI >= 25)

small_post_meal_data %>% # Filter for only control
  filter(Group == "CTR")


small_post_meal_data %>%
  filter(Group == "CTR" | BMI >= 25)
```




## Practice with filter and mutate

```{r}
# Task 1.
post_meal_data |>
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")

# Task 2.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```










## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename(
    p.glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data %>%
  select(id, BMI, Length, Weight, Age) %>%
  rename(Height = Length)
```
