---
title: "Reproducible Document"
author: "Oliver Bonde van Zwol"
format: html
---

```{r setup}
#| message: false
#| warning: false

library("tidyverse")

post_meal_data <- read.csv2(here::here("data/post-meal-insulin.csv"))


small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi) # make a new df with only the selected columns


tidier_post_meal_data <- small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) %>%
  filter(glykemi == 0) %>%
  select(-glykemi)

tidied_post_meal_data <- tidier_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```

## About me

-   Oliver
-   PhD

Just started on a **PhD** at *FGM*

## Simple code

```{r}
3 * 3
```




## Writing R code

2+2

```{r}
2 + 2
```

## Testing for Git

It is really cold outside


## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continous variable

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

See @fig-bmi-histo


## Exercise: discrete plots

```{r}
#| fig-cap: "Groups"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```


## Side by side plot
```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```


## Selecting  data

```{r}
select(post_meal_data, Age) # Selects only the Age column

select(post_meal_data, Age, BMI, Weight) # select only Age, BMI, and Weight column

select(post_meal_data, -Group) # removes the Group columns

select(post_meal_data, starts_with("PG")) # selects only columns starting with PG

select(post_meal_data, ends_with("DXA")) # selects all columns ending with DXA

select(post_meal_data, contains("0")) # selects all columns containing 0
```


## Renaming columns

```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```



## Filtering data


```{r}
small_post_meal_data %>%
  filter(glykemi == 1) # Filter for only glykemi of 1

small_post_meal_data %>%
  filter(BMI >= 25) # Filter for only 25 or greater BMI

small_post_meal_data %>% # Filter for only control
  filter(Group == "CTR")


small_post_meal_data %>%
  filter(Group == "CTR" | BMI >= 25)
```



## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename(
    p.glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data %>%
  select(id, BMI, Length, Weight, Age) %>%
  rename(Height = Length)
```


## modifying data

```{r}
tidier_post_meal_data %>%
  mutate(insulin_glucose_ratio = auc_insulin / auc_glucose)

tidier_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```


## Practice with filter and mutate

```{r}
# Task 1.
post_meal_data |>
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")

# Task 2..
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```


## Visualizing with a jitter plot

```{r}
tidied_post_meal_data %>%
  mutate(overweight = if_else(BMI >= 25, "overweight", "not overweight")) %>%
  ggplot(aes(x = overweight, y = insulin_glucose_ratio)) +
  geom_jitter() # within a ggplot it is not possible to pipe, thus + is needed
```

## Plotting three or more variables

```{r}
tidied_post_meal_data %>%
  filter(BMI < 30) %>%
  mutate(young = if_else(Age < 40, "young", "old")) %>%
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young)) + # rows facet in rows and cols = facet in columns
  labs(
    x = "Body Mass Index (BMI)",
    y = "Insulin to glucose ratio"
  )
```


## Make multi-dimensional plots!

```{r}
#| label: fig-multidimensional
#| fig-cap: "Two, mult-dimensional plots of our data."
#| fig-subcap:
#|   - "Distribution of percent body fat by group and age category."
#|   - "Relationship between BMI and bone mineral density, split by insulin sensitivity."
#| layout-ncol: 2
# Task 1.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100,
    young = if_else(Age < 40, "Young", "Old")
  ) |>
  ggplot(aes(x = Group, y = percent_body_fat, colour = young)) +
  geom_jitter(width = 0.2)

# Task 2.
post_meal_data |>
  mutate(
    insulin_glucose_ratio = (auc_ins / auc_pg),
    insulin_insensitive = if_else(insulin_glucose_ratio > mean(insulin_glucose_ratio), "Insensitive", "Sensitive")
  ) |>
  ggplot(aes(x = BMI, y = Bone.mineral.DXA, colour = insulin_insensitive)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(insulin_insensitive))
```


## Changing colours

```{r}
# Tasks 1. and 2.
post_meal_data |>
  mutate(hyperglycemia = if_else(glykemi == 1, "Yes", "No")) |>
  ggplot(aes(x = Group, fill = hyperglycemia)) +
  geom_bar(position = position_dodge()) +
  scale_fill_viridis_d(option = "magma")

# Tasks 3. and 4.
post_meal_data |>
  ggplot(aes(x = auc_pg, y = auc_ins, colour = BMI)) +
  geom_point() +
  scale_colour_viridis_c(option = "inferno")
```
